Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
More diagrams, working through comments

% APT: I don't think we need this.
%% \begin{figure}
%%   \begin{tabular}{| l | l | l |}
%%     \hline
%%     Set / & Names & Purpose \\
%%     Class & & \\
%%     \hline
%%     \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
%%     \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
%%     & {\tt a2} -- {\tt a7} & Caller-saved args \\
%%     & {\tt ra} & Return Address \\
%%     \hline
%%     \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
%%     \(\sealed\) & {\tt sp} & Stack Pointer \\
%%     \hline
%%     \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
%%     \(\public\) & {\tt tp} & Thread Pointer \\
%%     \hline
%%   \end{tabular}
%%   \caption{RISC-V integer register set}
%%   \label{fig:RISCVregs}
%% \end{figure}

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over the details of the target machine
architecture and ABI, the set of security-relevant operations, the rules that
govern evolution of the security context, the set of observable events, and
a notion of value compatibility.
\apt{Not quite true: the property definitions require a notion of
  call entry.}
\apt{Any more? Give a more listlike presentation?
  Need to clarify exactly what the parameters are, both in the initial
  presentation and in the various extensions, including the appendix.}

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
The details of value structure depend on the specific machine being modelled;
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.
Values are compatible if their non-payload information is identical, 
even though their payloads may differ.
\apt{This is still not very convincing!
  empting to say that only the payloads can affect the step function, but that is
  too strong, right? (OK for tags, but not, e.g. capabilities?}

@@

The security semantics operates in parallel with the 
Each state element (memory word or register) is given a \emph{security class} 
\(l \in \{\public, \object, \sealed, \unsealed\}\).
A \emph{view} \(V \in \mathit{VIEW}\) maps elements to security classes.
For any security class \(l\), we write \(l(V)\)
to denote the set of elements \(\component\) such that \(V ~ \component = l\).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
\apt{This doesn't quite match example, where there is magically an initial argument register labeled \(\active\).}

@@

Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
is the same. That is, we compare up to deletion of \(\tau\) events.
\apt{Didn't we use to say something about the (arguably) odd nature of this definition?}

>>>>

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over: (1) the details of  the target machine
architecture and ABI, (2) the set of observable events, and (3) a notion of value compatibility.
They can be extended to a more sophisticated machine model by expanding the set of
security-relevant events and defining

\apt{Any more? Give a more listlike presentation?
  Need to clarify exactly what the parameters are, both in the initial
  presentation and in the various extensions, including the appendix.}
\sna{Need to come back to this.}

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
The details of value structure depend on the specific machine being modelled;
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.
In the tagged setting, values are compatible if their tags are identical.
\apt{This is still not very convincing!
  empting to say that only the payloads can affect the step function, but that is
  too strong, right? (OK for tags, but not, e.g. capabilities?}
\sna{Yeah, I don't think we can say something universal about when values should be
  considered compatible.}

@@

The security semantics operates in parallel with the 
Each state element (memory word or register) is given a \emph{security class} 
\(l \in \{\public, \object, \sealed, \unsealed\}\).
A \emph{view} \(V \in \mathit{VIEW}\) maps elements to security classes.
For any security class \(l\), we write \(l(V)\)
to denote the set of elements \(\component\) such that \(V ~ \component = l\).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved except for those
that contain arguments at the start of execution, which are \(\object\), and \(\public\) otherwise.

@@

Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
is the same. That is, we compare up to deletion of \(\tau\) events.
Note that this results in an infinite silent trace being similar to
any trace. So, a trace that silently diverges due to a failstop will
fulfill this property vacuously.
